Accuracy and precision of zero-heat-flux temperature monitoring: A systematic review and meta-analysis

### 2019-12-03 manuscript/manuscript.Rmd
params to show text for word output

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
description: A systematic review and meta-analysis
author:
- name: Aaron Conway
  affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: aaron.conway@utoronto.ca
  url: https://www.aaronconway.info
- name: Megan Bittner
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Dan Phan
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Navpreet Kamboj
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: navpreet.kamboj@mail.utoronto.ca
- name: Kristina Chang
  affiliation: Toronto General Hospital, UHN
- name: Peter Collins
  affiliation: Toronto General Hospital, UHN
- name: Elizabeth Tipton
  affiliation: Dapartment of Statistics, Northwestern University
- name: Matteo Parotto
  affiliation: Toronto General Hospital, UHN
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: distill::distill_article
repository_url: https://github.com/awconway/zhf-review
twitter:
  creator: "@aw_conway"
---

@@

Zero-heat-flux thermometers are non-invasive devices that can be used to continuously monitor temperature. These devices are increasingly being used in clinical practice to substitute for invasive core temperature measurement during surgery and in critical care settings. The aim of this review was to determine the accuracy and precision of zero-heat-flux temperature measurements.

@@

Medline and EMBASE were searched for studies that reported on a measurement of core or peripheral temperature that coincided with a measurement from the zero-heat-flux device. Study selection and quality assessment (using the QUADAS-2) was performed independently. The GRADE approach was used to summarize the strength of the body of evidence.  Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (termed population limits of agreement) were calculated.

@@

Sixteen studies were included. The primary meta-analysis of zero-heat-flux versus core temperature measurements consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. Data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included. The pooled estimate for the mean bias was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C. The GRADE evidence quality rating was downgraded to moderate due to concerns about study limitations. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis.

@@

The range of uncertainty in the accuracy of the zero-heat-flux thermometer should be taken into account when using measurements from this device to inform clinical decision-making. Clinicians should therefore consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as 1°C higher or lower than core temperature. Use of this device may not be appropriate in situations where a difference in temperature of less than 1°C is important to detect.

@@

Published studies were found by searching Medline and EMBASE to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located [here](https://aconway.dev/zhf). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses.[@team2017r] All data and R code used in the meta-analyses can be located [here](https://github.com/awconway/zero-heat-flux-review).

@@

```{r fig-1, fig.align="center", fig.cap="PRISMA Flow Diagram", echo=FALSE, fig.width=5.5, fig.path=""}
# https://github.com/awconway/ggprisma
ggprisma::ggprisma(retrieved = 130, included = 16, duplicates = 35, full_text = 22, wrong_intervention = 4, wrong_comparator = 2, wrong_design = 1, awaiting_classification = 2)
```

The characteristics of included studies can be viewed [here](https://aconway.dev/zhf). The primary comparison of zero-heat-flux versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were judged as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Measurements` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants.

@@

```{r fig-2, preview=TRUE, layout="l-screen-inset", fig.align="center", fig.cap="Graphical overview of findings for the primary analysis", echo=FALSE, fig.path=""}
include_graphics("gofer_core.png")
```

@@

```{r table-1, layout="l-page-outset", echo=FALSE}
library(kableExtra)
  ma <- list(primary, low_risk, ICU, OT,  NPA,SL, conflict)
names(ma) <- c("Primary", "Low risk studies", "Intensive Care Unit", 
               "Surgery", "Nasopharyngeal", "Sublingual",
               "Studies not funded by industry")
df <- bind_rows(ma, .id="Comparison")
df %>% 
  rename(" " = Comparison, "Mean bias" = bias_mean, "Variance" = sd2_est, "Tau-squared" = tau_est, "Lower bound for 95% limit of agreement" = LOA_L, "Upper bound for 95% limits of agreement" = LOA_U, "Outer confidence interval for lower 95% limits of agreement" = CI_L_rve, "Outer confidence interval for upper 95% limits of agreement" = CI_U_rve) %>% 
    gt() %>%
  tab_header(
    title = "Table 1. Results of meta-analyses"  )  %>%
  tab_spanner(
    label = "Population limits of agreement",
    columns = vars(
      `Outer confidence interval for lower 95% limits of agreement`, `Outer confidence interval for upper 95% limits of agreement`)
  )
```

@@

>>>>

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring: A systematic review and meta-analysis"
author:
- name: Aaron Conway
  affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: aaron.conway@utoronto.ca
  url: https://www.aaronconway.info
- name: Megan Bittner
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Dan Phan
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
- name: Navpreet Kamboj
  affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
  email: navpreet.kamboj@mail.utoronto.ca
- name: Kristina Chang
  affiliation: Toronto General Hospital, UHN
- name: Peter Collins
  affiliation: Toronto General Hospital, UHN
- name: Elizabeth Tipton
  affiliation: Department of Statistics, Northwestern University
- name: Matteo Parotto
  affiliation: Toronto General Hospital, UHN
# date: "`r Sys.Date()`"
repository_url: https://github.com/awconway/zhf-review
twitter:
  creator: "@aw_conway"
bibliography: bibliography.bib
csl: anesthesiology.csl
output: word_document
# output: distill::distill_article
params:
  distill: FALSE
  word: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=FALSE,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r conditional_title_page, results='asis', include=params$word}
cat("
## Author information

<p>Aaron Conway PhD^1,2,3^
<p>Megan Bittner BSc^2^
<p>Dan Phan BSc^2^
<p>Navpreet Kamboj BScN^2^
<p>Kristina Chang MScN^4^
<p>Peter Collins MStatistics^4^
<p>Elizabeth Tipton PhD^5^
<p>Matteo Parotto MD^4^

<p>^1^ Peter Munk Cardiac Centre, University Health Network
<p>^2^Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
<p>^3^School of Nursing, Queensland University of Technology (QUT)
<p>^4^Toronto General Hospital, UHN
<p>^5^Department of Statistics, Northwestern University

## Corresponding author

Dr Aaron Conway BN (Hons) PhD
RBC Chair in Cardiovascular Nursing Research | Peter Munk Cardiac Centre | Toronto General Hospital | University Health Network
Clinician Investigator | Toronto General Hospital Research Institute | University Health Network
Assistant Professor | Lawrence S. Bloomberg Faculty of Nursing | University of Toronto

585 University Ave, 4NU - 476
Toronto, ON  M5G 2N2

e. aaron.conway@utoronto.ca  | t. (416) 946-7112

## Declarations

Clinical trial number: Not applicable

Prior presentations: Not applicable

Abstract word count: 300 words

Total word count: 3075 words

Introduction word count: 303 words

Discussion word count: 781 words

Number of figures: 2

Number of tables: 1

Abbreviated title: Accuracy of zero-heat-flux thermometers

Summary statement: Clinicians should consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as 1°C higher or lower than core temperature.

Funding statement: Support was provided solely from institutional and/or departmental sources.

Conflicts of interest: The authors declare no competing interests.

")
```

@@

Zero-heat-flux thermometers are non-invasive devices that can be used to continuously monitor temperature. These devices are increasingly being used to substitute for invasive core temperature measurement during surgery and in critical care. The aim of this review was to determine the accuracy and precision of zero-heat-flux temperature measurements.

@@

Medline and EMBASE were searched for studies that reported on a measurement of core or peripheral temperature that coincided with a measurement from the zero-heat-flux device. Study selection and quality assessment (using QUADAS-2) was performed independently. The GRADE approach was used to summarize the strength of the evidence.  Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (population limits of agreement) were calculated.

@@

Sixteen studies were included. The primary meta-analysis of zero-heat-flux versus core temperature consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. Data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included. The pooled estimate for the mean bias was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C. The GRADE evidence quality rating was downgraded to moderate due to concerns about study limitations. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis.

@@

The range of uncertainty in the accuracy of the zero-heat-flux thermometer should be taken into account when using this device to inform clinical decision-making. Clinicians should therefore consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as 1°C higher or lower than core temperature. Use of this device may not be appropriate in situations where a difference in temperature of less than 1°C is important to detect.

@@

Published studies were found by searching Medline and EMBASE to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located at this [link](https://aconway.dev/zhf). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses.[@team2017r] All data and R code used in the meta-analyses can be located at this [link](https://github.com/awconway/zhf-review).

@@

```{r fig-1, fig.align="center", fig.cap="PRISMA Flow Diagram", echo=FALSE, fig.width=5.5, fig.path="", include=params$distill}
# https://github.com/awconway/ggprisma
ggprisma::ggprisma(retrieved = 130, included = 16, duplicates = 35, full_text = 22, wrong_intervention = 4, wrong_comparator = 2, wrong_design = 1, awaiting_classification = 2)
```

The characteristics of included studies can be viewed at this [link](https://aconway.dev/zhf). The primary comparison of zero-heat-flux versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were judged as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Measurements` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants.

@@

```{r fig-2, preview=TRUE, layout="l-screen-inset", fig.align="center", fig.cap="Graphical overview of findings for the primary analysis", echo=FALSE, fig.path="", include=params$distill}
include_graphics("gofer_core.png")
```

@@

```{r table-1, layout="l-page-outset", include=params$distill}
library(kableExtra)
  ma <- list(primary, low_risk, ICU, OT,  NPA,SL, conflict)
names(ma) <- c("Primary", "Low risk studies", "Intensive Care Unit", 
               "Surgery", "Nasopharyngeal", "Sublingual",
               "Studies not funded by industry")
df <- bind_rows(ma, .id="Comparison")
df %>% 
  rename(" " = Comparison, "Mean bias" = bias_mean, "Variance" = sd2_est, "Tau-squared" = tau_est, "Lower bound for 95% limit of agreement" = LOA_L, "Upper bound for 95% limits of agreement" = LOA_U, "Outer confidence interval for lower 95% limits of agreement" = CI_L_rve, "Outer confidence interval for upper 95% limits of agreement" = CI_U_rve) %>% 
    gt() %>%
  tab_header(
    title = "Table 1. Results of meta-analyses"  )  %>%
  tab_spanner(
    label = "Population limits of agreement",
    columns = vars(
      `Outer confidence interval for lower 95% limits of agreement`, `Outer confidence interval for upper 95% limits of agreement`)
  )
```

@@

```{r  conditional_references, include=params$word, results='asis'}
cat("# References")
```
<div id="refs"></div>

```{r  conditional_figure_legend, include=params$word, results='asis'}
cat("# Figure legend

Figure 1: PRISMA Flow Diagram

Figure 2: Graphical overview of findings for the primary analysis")
```
